### 2017-11-02 submission/manuscript.Rmd
add last paragraph

- In closing to get at these more difficult questions the Power calculations and effect size in future studies need to be accounted for especially for adenoma where the predicted effect sizes are going to be small

>>>>

Although there are still questions that need to be answered for the microbiota and carcinoma a clearer framework is beginning to develop as to how this occurs. The role of the microbiota in adenoma is still not clear and part of the reason may be because many studies are not powered effectely to observe the small changes reported here. It is realistic to suspect that many changes in carcinoma could easily result in effect sizes that are 30% or more between the case and control. Most of the studies have sufficient power to detect these changes. In contrast, our data suggests that the adenoma effect size is relatively small. None of the studies analyzed  were properly powered to detect a 10% or lower change between case and controls and this may well be the range in which differences occur for adenoma. Future studies investigating adenoma and the microbiota need to take these factors into consideration if we are to work out the role of the microbiota in adenoma genesis.

### 2017-11-02 submission/manuscript.tex
add last paragraph

\begin{itemize}
\tightlist
\item
  In closing to get at these more difficult questions the Power
  calculations and effect size in future studies need to be accounted
  for especially for adenoma where the predicted effect sizes are going
  to be small
\end{itemize}

@@

\textbf{\emph{Sequence Processing:}} For the majority of studies raw
sequences were downloaded from the SRA
(\url{ftp://ftp-trace.ncbi.nih.gov/sra/sra-instant/reads/ByStudy/sra/SRP/})
and metadata was obtained from the following website:
\url{http://www.ncbi.nlm.nih.gov/Traces/study/} by searching the
respective accession number of the study. Of the studies that did not
have sequences and meta data on the SRA one study had the data stored on
DBGap {[}4{]} and four studies the data was obtained directly from the
authors {[}10,12,15,17{]}. Each study was processed using the mothur
(v1.39.3) software program {[}20{]}. Where possible quality filtering
utilized the default methods used in mothur for either 454 or Illumina
based sequencing. If it was not possible to use these defaults the
author stated quality cut-offs were used instead. Chimeras were
identifed and removed using the VSEARCH {[}{\textbf{???}}{]} program and
\emph{de novo} OTU clustering at 97\% similarity using the OptiClust
algorithm {[}21{]} was utilized.

@@

\hypertarget{ref-westcott_opticlust_2017}{}
21. Westcott SL, Schloss PD. OptiClust, an Improved Method for Assigning
Amplicon-Based Sequence Data to Operational Taxonomic Units. mSphere.
2017;2.

>>>>

Although there are still questions that need to be answered for the
microbiota and carcinoma a clearer framework is beginning to develop as
to how this occurs. The role of the microbiota in adenoma is still not
clear and part of the reason may be because many studies are not powered
effectely to observe the small changes reported here. It is realistic to
suspect that many changes in carcinoma could easily result in effect
sizes that are 30\% or more between the case and control. Most of the
studies have sufficient power to detect these changes. In contrast, our
data suggests that the adenoma effect size is relatively small. None of
the studies analyzed were properly powered to detect a 10\% or lower
change between case and controls and this may well be the range in which
differences occur for adenoma. Future studies investigating adenoma and
the microbiota need to take these factors into consideration if we are
to work out the role of the microbiota in adenoma genesis.

@@

\textbf{\emph{Sequence Processing:}} For the majority of studies raw
sequences were downloaded from the SRA
(\url{ftp://ftp-trace.ncbi.nih.gov/sra/sra-instant/reads/ByStudy/sra/SRP/})
and metadata was obtained from the following website:
\url{http://www.ncbi.nlm.nih.gov/Traces/study/} by searching the
respective accession number of the study. Of the studies that did not
have sequences and meta data on the SRA one study had the data stored on
DBGap {[}4{]} and four studies the data was obtained directly from the
authors {[}10,12,15,17{]}. Each study was processed using the mothur
(v1.39.3) software program {[}20{]}. Where possible quality filtering
utilized the default methods used in mothur for either 454 or Illumina
based sequencing. If it was not possible to use these defaults the
author stated quality cut-offs were used instead. Chimeras were
identifed and removed using the VSEARCH {[}21{]} program and \emph{de
novo} OTU clustering at 97\% similarity using the OptiClust algorithm
{[}22{]} was utilized.

@@

\hypertarget{ref-rognes_vsearch_2016}{}
21. Rognes T, Flouri T, Nichols B, Quince C, Mahé F. VSEARCH: A
versatile open source tool for metagenomics. PeerJ. 2016;4:e2584.

\hypertarget{ref-westcott_opticlust_2017}{}
22. Westcott SL, Schloss PD. OptiClust, an Improved Method for Assigning
Amplicon-Based Sequence Data to Operational Taxonomic Units. mSphere.
2017;2.
